---
title: "Research Update"
author: "Tyler Hoecker"
date: "11/13/2018"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(lubridate)
library(lme4)
```

```{r, include = F}
source('code/read_soil_data.R')
soil_sum <- soil_df %>% 
  filter(time >= as.POSIXct(c("2018-06-20 00:00:00")) & time <= as.POSIXct(c("2018-08-20 00:00:00"))) %>% 
  group_by(site, aspect, time, variable) %>%
  summarise(value = mean(value)) %>% 
  spread(variable, value) %>% 
  group_by(site, aspect) %>% 
  summarise_at(vars(mois, temp),
               funs(low = quantile(., 0.05), 
                    med = quantile(., 0.50), 
                    high = quantile(., 0.95))
               , na.rm = T) 

# ATMOS data
source('code/read_atmos_data.R') #creats 'atmos_df'
atmos_sum <- atmos_df %>% 
  filter(time >= as.POSIXct(c("2018-06-20 00:00:00")) & time <= as.POSIXct(c("2018-08-20 00:00:00")),
         aspect != 'Grizz') %>% 
  group_by(site, aspect) %>%
  summarise_at(vars(air_temp:solar),
               funs(low = quantile(., 0.05), 
                     med = quantile(., 0.50), 
                     high = quantile(., 0.95)), na.rm = T) %>% 
  mutate(solar_high = replace(solar_high, solar_high > 900 | solar_high < 840, NA)) 
  

# Import seedling data
source('code/read_seedling_data.R')


# Surival and germination figuring-------------------------------------------
# Germinated
germination <- seedlings %>% 
  filter(variable == 'height', species != 'control') %>% 
  select(site, aspect, species, frameID, cell, value) %>% 
  mutate(germinated = if_else(is.na(value), 0, 1)) %>% 
  unite(frameID, cell, col = 'uniqID', remove = FALSE) %>%
  group_by(site, aspect, species, frameID, uniqID) %>% 
  summarise(germinated = max(germinated)) %>% 
  group_by(site, aspect, species, frameID) #%>% 
  #mutate(problem = if_else(sum(germinated) == 0, 'Yes','No')) %>% 
  #filter(problem == 'No') 
  

# Survived
survival <- seedlings %>%
  filter(variable == 'height', species != 'control') %>% 
  mutate(survived = if_else(date > as.POSIXct("2018-10-01") & value > 0, 1, 0)) %>% 
  select(site, aspect, species, frameID, cell, survived) %>% 
  unite(frameID, cell, col = 'uniqID') %>%
  group_by(site, aspect, species, uniqID) %>% 
  summarise(survived = max(survived, na.rm = T))  
  

establishment_df <- full_join(germination, survival) %>% 
  ungroup()

estb_abio_df <- establishment_df %>% 
  full_join(atmos_sum, by = c('site','aspect')) %>% 
  full_join(soil_sum, by = c('site','aspect')) %>% 
  select(site, aspect, species, frameID, uniqID, germinated, survived,
         air_temp_med, solar_high, solar_med, atms_press_med,
         temp_high, temp_med, temp_low, 
         mois_high, mois_med, mois_low, 
         rel_hum_med, rel_hum_high) 
```

### New plots of germination and survival by aspect

#### Germination
```{r echo = FALSE, fig.dim = c(7, 5)}
colVals <- c('Flat' = '#009E73','North' = '#0072B2','South' = '#E69F00')

ggplot(estb_abio_df) +
  geom_col(aes(x = aspect, y = germinated, fill = aspect)) +
  #geom_errorbar(aes(x = aspect, ymin = low, ymax = high)) +
  #geom_boxplot(aes(x = aspect, y = count, fill = aspect)) +
  #scale_y_continuous(sec.axis = sec_axis(~./(50*5*4), name = "Percent seeds germinated")) +
  scale_fill_manual(values = colVals, name = 'aspect') +
  #scale_y_continuous(limits = c(0,150), breaks = seq(0,150,50)) +
  labs(x = 'Aspect', y = 'Percent germinants survived') +  
  facet_wrap(~species, ncol = 1, scales = 'free_y') +
  theme_bw(base_size = 14) +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

#### Survival
```{r echo = FALSE, fig.dim = c(7, 5)}
colVals <- c('Flat' = '#009E73','North' = '#0072B2','South' = '#E69F00')

ggplot(estb_abio_df) +
  geom_col(aes(x = aspect, y = survived, fill = aspect)) +
  #geom_errorbar(aes(x = aspect, ymin = low, ymax = high)) +
  #geom_boxplot(aes(x = aspect, y = count, fill = aspect)) +
  #scale_y_continuous(sec.axis = sec_axis(~./(50*5*4), name = "Percent seeds germinated")) +
  scale_fill_manual(values = colVals, name = 'aspect') +
  #scale_y_continuous(limits = c(0,150), breaks = seq(0,150,50)) +
  labs(x = 'Aspect', y = 'Percent germinants survived') +  
  facet_wrap(~species, ncol = 1, scales = 'free_y') +
  theme_bw(base_size = 14) +
  theme(strip.background = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```




### Exploring binary outcomes as a function of environmental variables  

#### Bar plots of **germination** by environmental variable:  
```{r echo = FALSE, fig.dim = c(12, 6)}
temp <- estb_abio_df %>%
  gather(variable, value, -site, -aspect, -species, -frameID, -uniqID, -germinated, -survived) 
  
ggplot(temp, aes(factor(germinated), y = value, fill=factor(germinated))) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free_y", nrow = 2) +
  scale_fill_grey(start = 0.8, end = 0.4) +
  theme_bw(base_size = 12) 
```


#### Bar plots of **survival** by environmental variable:  
```{r echo = FALSE, fig.dim = c(12, 6)}
temp <- estb_abio_df %>%
  gather(variable, value, -site, -aspect, -species, -frameID, -uniqID, -germinated, -survived) 
  
ggplot(temp, aes(factor(survived), y = value, fill=factor(survived))) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free_y", nrow = 2) +
  scale_fill_grey(start = 0.8, end = 0.4) +
  theme_bw(base_size = 12) 
```


### Modeling binary outcomes as a function of environmental variables

#### GLM with binomial (logit) link for germination
```{r}
m <- glm(germinated ~ mois_med + temp_med + air_temp_med + solar_high + rel_hum_high + 
           as.factor(species), 
           data = as.data.frame(estb_abio_df), family = 'binomial')
summary(m)
confint(m)
exp(coef(m))
```


#### GLM with binomial (logit) link for survival
```{r}
m <- glm(survived ~ mois_med + temp_med + air_temp_med + solar_high + rel_hum_high + 
           as.factor(species), 
           data = as.data.frame(estb_abio_df), family = 'binomial')
summary(m)
confint(m)
exp(coef(m))
```

#### Mixed-effects model with binomial (logit) link for germination
Soil moisture, soil temp., air temp., 95th solar radiation, 95th rel. humidity are fixed effects. Species and site are random effects.
```{r}
m <- glmer(germinated ~ mois_med + temp_med + air_temp_med + solar_high + rel_hum_high + 
             (1 | species) + (1 | site), 
           data = as.data.frame(estb_abio_df), family = 'binomial', nAGQ=0)
summary(m)
```

#### Mixed-effects model with binomial (logit) link for survival
Soil moisture, soil temp., air temp., 95th solar radiation, 95th rel. humidity are fixed effects. Species and site are random effects.
```{r}
m <- glmer(survived ~ mois_med + temp_med + air_temp_med + solar_high + rel_hum_high + 
             (1 | species) + (1 | site), 
           data = as.data.frame(estb_abio_df), family = 'binomial', nAGQ=0)
summary(m)
```
